---
title: "Quality Control for bulk RNA-seq data"
author: "Ximing Ran"
date: "2025-05-15"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: false
  html_document:
    # code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js

---

```{r setup, include=FALSE}
# load libraries
library(tidyverse)
library(knitr)

knitr::opts_chunk$set(
  # echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.width = 9,
  fig.height = 4,
  fig.path = "../results/01-QC/"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
knitr::kable(head(mtcars[, 1:4]), "simple")
```

```{r}
library(tibble)
library(tidyr)
library(dplyr)
library(rtracklayer)
```



```{r local_function_load}
# load function from local files

```

# 1. Read the count data     
In this section, we will read the raw count data from the neuron_bulkRNA folder. 


```{r load_bulkRNA_data}
sample_df <- read.csv(here::here("data", "neuron_bulkRNA",
                                  "sample_info.csv"), 
                       header = TRUE)
colnames(sample_df) <- c("Raw", "SampleID", "Condition")


condition_list <- sample_df$Condition

sample_list <- sample_df$Raw
sample_ID_list <- sample_df$SampleID
ref_sample <- paste0(sample_list[1], "_final_gene_with_names.csv")
counts_ref <- read.csv(here::here("data", "neuron_bulkRNA",
                                  "raw",ref_sample), 
                       header = TRUE)

counts <- tibble()
for (i in 1:length(sample_list)) {
  sample_raw <- sample_list[i]
  gene_file <- here::here("data", "neuron_bulkRNA","raw",
                          paste0(sample_raw, "_final_gene_with_names.csv"))
  gene <- read.csv(gene_file, header = TRUE)
  
  if (i == 1) {
    counts <- gene[, c("GeneID", "GeneName")]
    counts <- counts %>% mutate(!!sample_ID_list[i] := gene$Count)
  } else {
    temp_counts <- gene[, c("GeneID", "Count")]
    colnames(temp_counts)[2] <- sample_ID_list[i]
    counts <- merge(counts, temp_counts, by = "GeneID")
  }
}

# replace all - in column names with _
colnames(counts) <- gsub("-", "_", colnames(counts))
write.csv(counts,here::here("data", "neuron_bulkRNA", 
                            "neuron_bulkRNA_counts_raw.csv"),
          row.names = FALSE)
```




# 2. Map the gene name.    

In this section, we will map the gene name to the gene ID using the GTF file. Since some gene may have different gene name, we will check if the gene name is unique. If not, we will find the gene that has more than one gene name. And merge their counts. The gene annotation comes from the the file `Homo_sapiens.GRCh38.106.gtf`. The final table will be stored in `results/01-QC/synaptosomes_bulkRNA_counts_cleaned.csv`. 


```{r Map_gene_name}

# read the gtf file
gtf_data <- import(here::here("data", "ref","Homo_sapiens.GRCh38.106.gtf"))

gtf_df <- as.data.frame(gtf_data)

gtf_genes <- gtf_df %>% 
  filter(type == "gene") %>% 
  select(gene_id, gene_name) 
colnames(gtf_genes) <- c("GeneID","GeneName")

# check the annotation format
head(gtf_genes)
gtf_gene <- na.omit(gtf_genes)

# check if the gene is unique
unique_genename <- length(unique(gtf_genes$GeneName)) == nrow(gtf_genes)
print(paste("GeneName is unique:", unique_genename))


# Find the genes that appear more than once
duplicate_genes <- gtf_genes$GeneName[duplicated(gtf_genes$GeneName)]
duplicate_genes <- unique(duplicate_genes)

# Correct sprintf statement
print(sprintf("There are %d genes with duplicate gene names", 
              length(duplicate_genes)))

# Merge counts data with GTF information
merged_data <- counts %>% 
  left_join(gtf_genes, by = "GeneID")%>%
  select(-GeneName.x,-GeneID)

names(merged_data)[names(merged_data) == "GeneName.y"] <- "gene"

# find the same gene
aggregated_data <- merged_data %>%
  group_by(gene) %>%
  summarise(across(everything(), \(x) sum(x, na.rm = TRUE)))



# Using sprintf (Recommended for better formatting)
print(sprintf("The shape of the count matrix is: %d x %d", 
              dim(aggregated_data)[1], dim(aggregated_data)[2]))

# clean the NA in gene
aggregated_data <- aggregated_data[complete.cases(aggregated_data), ]
# check if here is any NA in gene
print(sprintf("There are %d genes with NA gene name", 
              sum(is.na(aggregated_data$gene))))

write.csv(aggregated_data, here::here("data", "neuron_bulkRNA", 
                                      "neuron_bulkRNA_counts_cleaned.csv"), 
          row.names = FALSE)
```


# 3. make gene refernce 
ALso, we will make a reference table for the gene length. The reference table will be stored in `results/01-QC/gene_lengths.csv`. 


```{r make_ref_for_gene}
# Filter for rows with gene information only
genes_df <- gtf_df %>%
  filter(type == "gene") %>%
  select(seqnames, start, end, gene_id, gene_name)


genes_df <- genes_df %>%
  mutate(gene_length = end - start + 1) %>%
  select(gene_id, gene_name, gene_length)

# Save to CSV
write.csv(genes_df,here::here("data", "ref", "gene_lengths.csv"), 
          row.names = FALSE)

```


\newpage
```{r}
sessionInfo()
```

