---
title: "WGCNA"
subtitle: "Neuron"
author: "Ximing Ran"
date: "2025-05-15"
output:
  pdf_document:
    toc: true
    toc_depth: 3
    number_sections: false
  html_document:
    # code_folding: hide
    toc: true
    toc_float: true
hitheme: tomorrow
highlighter: highlight.js

---

```{r setup, include=FALSE}
# load libraries
library(tidyverse)
library(knitr)
set.seed(2025)

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.path = "./results/03-WGCNA/Analysis_figure/"
)

theme_set(theme_bw() + theme(legend.position = "bottom"))
knitr::kable(head(mtcars[, 1:4]), "simple")
```

```{r}
library(tibble)
library(tidyr)
library(dplyr)
library(rtracklayer)
library(WGCNA)
```



```{r local_function_load}
# load function from local files
source(here::here("source", "DEG_functions.R"))
```

\newpage

# 1. Read the count data
In this section, we will read the clean count data from the synaptosomes_miRNARNA folder. The data is stored in the format of `.csv`. We will read the data and filter the columns based on the condition list. The final table will be stored in `results/02-DEG-Vehicle/DESeq2_results.csv`.

```{r load_bulkRNA_data}
input_count <- read.csv(here::here("data", "neuron_bulkRNA",
                                      "neuron_bulkRNA_counts_cleaned.csv"))
counts <-  as.data.frame(input_count) %>% 
  column_to_rownames(var = "gene")
colnames(counts) <- gsub("-", "_", colnames(counts))


# raw sample list
sample_list_raw <- read.csv(here::here("data", "neuron_bulkRNA",
                                      "sample_info.csv")) %>%
                    mutate(condition = Diagnosis) 
colnames(sample_list_raw) <- c("Raw", "SampleID", "Diagnosis","Condition")
# order the sample list by Condition
sample_list_raw <- arrange(sample_list_raw,  desc(Condition))

sample_list_raw <- sample_list_raw %>% mutate(SampleID = gsub("-", "_", SampleID))

# Ensure the column names of counts exist in Sample.name
new_colnames <- sample_list_raw$SampleID[match(colnames(counts), sample_list_raw$SampleID)]

# Assign new column names
colnames(counts) <- new_colnames


# sort the columns by the colname
condition_list <- data.frame(
  group = sample_list_raw$Condition
)

row.names(condition_list) <- sample_list_raw$SampleID

counts<- counts[, rownames(condition_list)]

# load the 22q gene_list
target_22q_gene <- read.csv(here::here("data","ref" ,"22q_gene_2024_10_17.csv"))
target_gene <- target_22q_gene$gene
target_gene <- target_gene[1:(length(target_gene) - 4)]

gene_name_mapping<- readRDS(here::here("data","ref" ,"gene_name_mapping.rds"))



# load the DEG with logFC 1.0
DEG_1.0 <- read.csv("results/02-DEG/02-DEG/02_DEG_log2fc_1.csv")
# counts <- counts[rownames(counts) %in% DEG_1.0 $gene, ]

```


# 2. Normalization to RPKMs

```{r normalization}
library(countToFPKM)


count_input <-  counts
count_input <- count_input[rowSums(count_input) > 0, ]

gene_length <- read.csv(here::here("data", "ref",
                                   "gene_lengths.csv"))
# filter the rows with gene names not empty
gene_length <- gene_length %>% filter(!is.na(gene_name))




# 1. Match gene lengths to the rows of count_input
idx <- match(rownames(count_input), gene_length$gene_name)
if (any(is.na(idx))) {
  stop("Some genes in count_input have no matching entry in gene_length")
}
length_kb <- gene_length$gene_length[idx] / 1000  # bases to kilobases

# 2. Compute library sizes in millions of reads
lib_sizes <- colSums(count_input)
lib_size_million <- lib_sizes / 1e6

# 3. Compute FPKM
#    First divide counts by library size (in millions), then divide by gene length (in kb)
fpkm_step1 <- sweep(count_input, 2, lib_size_million, FUN = "/")
fpkm     <- sweep(fpkm_step1,    1, length_kb,       FUN = "/")




```



# 2. WGCNA analysis




```{r find_power, fig.width=8, fig.height=4}

library(WGCNA)
allowWGCNAThreads()          # allow multi-threading (optional)
#> Allowing multi-threading with up to 4 threads.

# Choose a set of soft-thresholding powers
powers = c(c(1:10), seq(from = 12, to = 20, by = 2))


input_mat = t(fpkm)
# Call the network topology analysis function
sft = pickSoftThreshold(
  input_mat ,             # <= Input data
  # blockSize = 30,
  powerVector = powers,
  verbose = 5
  )


par(mfrow = c(1,2));
cex1 = 0.9;

plot(sft$fitIndices[, 1],
     -sign(sft$fitIndices[, 3]) * sft$fitIndices[, 2],
     xlab = "Soft Threshold (power)",
     ylab = "Scale Free Topology Model Fit, signed R^2",
     main = paste("Scale independence")
)
text(sft$fitIndices[, 1],
     -sign(sft$fitIndices[, 3]) * sft$fitIndices[, 2],
     labels = powers, cex = cex1, col = "red"
)
abline(h = 0.6, col = "red")
plot(sft$fitIndices[, 1],
     sft$fitIndices[, 5],
     xlab = "Soft Threshold (power)",
     ylab = "Mean Connectivity",
     type = "n",
     main = paste("Mean connectivity")
)
text(sft$fitIndices[, 1],
     sft$fitIndices[, 5],
     labels = powers,
     cex = cex1, col = "red")


```


```{r build_network}
picked_power = 7
temp_cor <- cor       
cor <- WGCNA::cor         # Force it to use WGCNA cor function (fix a namespace conflict issue)
netwk <- blockwiseModules(input_mat,                # <= input here

                          # == Adjacency Function ==
                          power = picked_power,                # <= power here
                          networkType = "signed",

                          # == Tree and Block Options ==
                          # deepSplit = 2,
                          # pamRespectsDendro = F,
                          # detectCutHeight = 0.75,
                          minModuleSize = 30,
                          # maxBlockSize = 4000,

                          # == Module Adjustments ==
                          # reassignThreshold = 0,
                          # mergeCutHeight = 0.25,

                          # == TOM == Archive the run results in TOM file (saves time)
                          saveTOMs = T,
                          saveTOMFileBase = "ER",

                          # == Output Options
                          numericLabels = T,
                          verbose = 3)
```

```{r plot_dendrogram, fig.width=8, fig.height=4}
# Convert labels to colors for plotting
mergedColors = labels2colors(netwk$colors)

# Plot the dendrogram and the module colors underneath
# …before plotting, load your libraries, etc.
library(WGCNA)

# create a transparent‐bg PDF via Cairo
cairo_pdf(
  filename = file.path("results", "03-WGCNA", "1.dendrogram.pdf"),
  width    = 15,
  height   = 8,
  bg       = "transparent"
)

# make sure the plotting region itself is transparent
par(bg = NA)

# your dendrogram + module‐color strip
plotDendroAndColors(
  netwk$dendrograms[[1]],
  mergedColors[netwk$blockGenes[[1]]],
  "Module colors",
  dendroLabels = FALSE,
  hang         = 0.03,
  addGuide     = TRUE,
  guideHang    = 0.05
)

dev.off()



#––– PNG with transparent background (raster) –––#
png(
  filename = file.path("results", "03-WGCNA", "1.dendrogram.png"),
  width    = 15,
  height   = 8,
  units    = "in",       # inches
  res      = 300,        # 300 DPI
  bg       = "transparent"
)
par(bg = NA)
plotDendroAndColors(
  netwk$dendrograms[[1]],
  mergedColors[netwk$blockGenes[[1]]],
  "Module colors",
  dendroLabels = FALSE,
  hang         = 0.03,
  addGuide     = TRUE,
  guideHang    = 0.05
)
dev.off()

```



```{r module_eigengenes}
MEs0 <- moduleEigengenes(input_mat, mergedColors)$eigengenes

table(mergedColors)
length(unique(mergedColors))

# save the genes for each module
module_genes <- data.frame(
  gene = colnames(input_mat),
  module = mergedColors
)


# sort the genes by module
module_genes <- module_genes %>% arrange(module)
# save the module genes
write.csv(module_genes, file = "results/03-WGCNA/module_genes.csv", row.names = F)



# do the module eigengene analysis
moduleTraitCor = cor(MEs0, condition_list, use = "p")
dim(condition_list)




# 1. Combine MEs and group labels into one data.frame
library(tibble)
library(dplyr)
library(tidyr)
library(broom)

# assume MEs0 is your eigengene matrix and condition_list is a data.frame with rownames sample IDs
dat <- MEs0 %>%
  rownames_to_column("sample") %>%                # bring sample IDs into a column
  as_tibble() %>%
  left_join(
    condition_list %>% rownames_to_column("sample"),
    by = "sample"
  ) %>%
  pivot_longer(
    cols = starts_with("ME"),
    names_to = "module",
    values_to = "eigengene"
  )

# 2. Perform a two‐sample t‐test for each module
module_tests <- dat %>%
  group_by(module) %>%
  do(tidy(t.test(eigengene ~ group, data = .))) %>%
  ungroup()

# 3. Adjust p‐values across modules (e.g. FDR)
module_tests <- module_tests %>%
  mutate(p.adj = p.adjust(p.value, method = "fdr")) %>%
  arrange(p.adj)

# 4. Inspect
module_tests %>% dplyr::select(module, estimate1, estimate2, statistic, p.value, p.adj)

# save the results
write.csv(module_tests, file = "results/03-WGCNA/module_trait_difference.csv", row.names = F)



```




```{r module_trait_correlation}
# 1. Prepare trait as numeric (0 = CTRL_Vehicle, 1 = 22q_Vehicle)
trait <- ifelse(condition_list$group == "22q", 1, 0)

# 2. Compute Pearson correlation and p-value for each module
corr_list <- lapply(
  X   = colnames(MEs0),
  FUN = function(mod) {
    ct <- cor.test(MEs0[, mod], trait, method = "pearson")
    data.frame(
      module    = mod,
      cor       = unname(ct$estimate),
      p.value   = ct$p.value
    )
  }
)
corr_df <- do.call(rbind, corr_list)

# 3. Adjust p-values for multiple testing
corr_df <- corr_df %>%
  arrange(p.value) %>%
  mutate(p.adj = p.adjust(p.value, method = "fdr"))


```


```{r module_trait_correlation2}
library(WGCNA)
# wrap trait in a data.frame
traits <- data.frame(group = trait)

# compute module–trait correlations and p-values
moduleTraitCor    <- cor(MEs0, traits, use = "p")
moduleTraitPvalue <- corPvalueStudent(moduleTraitCor, nSamples = nrow(MEs0))

# turn into a tidy table
corr_df2 <- data.frame(
  module  = rownames(moduleTraitCor),
  cor     = as.numeric(moduleTraitCor),
  p.value = as.numeric(moduleTraitPvalue)
) %>%
  arrange(p.value) %>%
  mutate(p.adj = p.adjust(p.value, method = "fdr"))

# inspect
corr_df2
# save the results

write.csv(corr_df2, file = "results/03-WGCNA/module_trait_correlation.csv", row.names = F)
```



```{r heatmap, fig.width=6, fig.height=12}
# # 1. Compute module–trait correlations and p‐values (if you haven't already)
# library(WGCNA)
# # prepare trait as numeric: 0 = CTRL_Vehicle, 1 = 22q_Vehicle
# traitNum <- ifelse(condition_list$group == "22q", 1, 0)
# traits   <- data.frame(Phenotype = traitNum)
# 
# # calculate correlation and Student’s p‐value
# moduleTraitCor    <- cor(MEs0, traits, use = "p")
# moduleTraitPvalue <- corPvalueStudent(moduleTraitCor, nSamples = nrow(MEs0))
# 
# # 2. Prepare a text matrix of “correlation (p‐value)”
# textMatrix <- paste0(
#   signif(moduleTraitCor, 2),
#   "(",
#   signif(moduleTraitPvalue, 1),
#   ")"
# )
# dim(textMatrix) <- dim(moduleTraitCor)
# rownames(textMatrix) <- rownames(moduleTraitCor)
# colnames(textMatrix) <- colnames(moduleTraitCor)
# 
# # 3. Draw the heatmap
# # labeledHeatmap(
# #   Matrix       = moduleTraitCor,
# #   xLabels      = colnames(traits),
# #   yLabels      = rownames(moduleTraitCor),
# #   colorLabels  = FALSE,
# #   colors       = colorRampPalette(c("blue", "white", "red"))(50),
# #   textMatrix   = textMatrix,
# #   setStdMargins= FALSE,
# #   cex.text     = 0.6,
# #   zlim         = c(-1, 1),
# #   main         = "Module–Trait Correlations\n(cor, p value)"
# # )
# 
# 
# # —— after computing moduleTraitCor & moduleTraitPvalue and building textMatrix ——
# 
# # 1. Get p-values for your single trait column
# pv <- moduleTraitPvalue[, 1]
# 
# # 2. Determine module order by increasing p-value
# ord <- order(pv)
# 
# # 3. Reorder all matrices
# moduleTraitCor  <- moduleTraitCor[ord, , drop = FALSE]
# moduleTraitPvalue <- moduleTraitPvalue[ord, , drop = FALSE]
# textMatrix      <- textMatrix[ord, , drop = FALSE]
# 
# # —— now draw the heatmap with modules sorted by p-value ——
# 
# # labeledHeatmap(
# #   Matrix        = moduleTraitCor,
# #   xLabels       = colnames(traits),
# #   yLabels       = rownames(moduleTraitCor),
# #   colorLabels   = FALSE,
# #   colors        = colorRampPalette(c("blue", "white", "red"))(50),
# #   textMatrix    = textMatrix,
# #   setStdMargins = FALSE,
# #   cex.text      = 0.6,
# #   zlim          = c(-1, 1),
# #   main          = "Module–Trait Correlations\n(sorted by p-value)"
# # )



# 1. Compute module–trait correlations & p-values
library(WGCNA)
traitNum          <- ifelse(condition_list$group == "22q", 1, 0)
traits            <- data.frame(Phenotype = traitNum)
moduleTraitCor    <- cor(MEs0, traits, use = "p")
moduleTraitPvalue <- corPvalueStudent(moduleTraitCor, nSamples = nrow(MEs0))

# 2. Extract the p-values & compute stars
pv    <- moduleTraitPvalue[, 1]
stars <- ifelse(pv < 0.001, "***",
         ifelse(pv < 0.01,  "**",
         ifelse(pv < 0.05,  "*", "")))

# 3. Order modules by ascending p-value
ord <- order(pv)
moduleTraitCor    <- moduleTraitCor[ord, , drop = FALSE]
moduleTraitPvalue <- moduleTraitPvalue[ord, , drop = FALSE]
stars             <- stars[ord]

# 4. Build textMatrix: “correlation\n(p= p-value)stars”
textMatrix <- matrix(
  paste0(
    signif(moduleTraitCor, 2), "",
    "(p=", signif(moduleTraitPvalue, 1), ")",
    stars
  ),
  nrow = nrow(moduleTraitCor),
  ncol = ncol(moduleTraitCor),
  dimnames = dimnames(moduleTraitCor)
)

# 5a. Plot with labeledHeatmap (from WGCNA)
labeledHeatmap(
  Matrix        = moduleTraitCor,
  xLabels       = colnames(traits),
  yLabels       = rownames(moduleTraitCor),
  colorLabels   = FALSE,
  colors        = colorRampPalette(c("blue", "white", "red"))(50),
  textMatrix    = textMatrix,
  setStdMargins = FALSE,
  cex.text      = 0.6,
  zlim          = c(-1, 1),
  main          = "Module–Trait Correlations\n(cor, p-value, stars)"
)

# 5b. Or, using pheatmap
library(pheatmap)
pheatmap(
  moduleTraitCor,
  color           = colorRampPalette(c("blue","white","red"))(100),
  breaks          = seq(-1, 1, length.out = 101),
  display_numbers = textMatrix,
  main            = "Module–Trait Correlations\n(cor, p-value, stars)",
  cluster_rows    = FALSE,
  cluster_cols    = FALSE
)





library(pheatmap)

# 1. pull out module names + colours from your ME rownames
moduleNames  <- sub("^ME", "", rownames(moduleTraitCor))
moduleColors <- moduleNames   # each entry is e.g. "blue","turquoise",…

# 2. make a one‐column data.frame for the row‐annotations
annotation_row <- data.frame(Module = factor(moduleColors, levels = unique(moduleColors)))
rownames(annotation_row) <- moduleNames

# 3. build a named list telling pheatmap how to colour your Module factor
ann_colors <- list(
  Module = setNames(unique(moduleColors), unique(moduleColors))
)

# 4. draw!
pheatmap(
  moduleTraitCor,
  color              = colorRampPalette(c("blue","white","red"))(50),
  cluster_rows       = FALSE,
  cluster_cols       = FALSE,
  display_numbers    = textMatrix,
  fontsize_number    = 6,
  labels_row         = rownames(moduleTraitCor),
  labels_col         = colnames(traits),
  annotation_row     = annotation_row,
  annotation_colors  = ann_colors,
  annotation_legend  = FALSE,   # turn off module‐colour legend
  legend             = TRUE,    # keep the heatmap colour legend
  border_color       = "black",      # hide cell borders
number_color = "black", # text colour for the numbers
  main               = "Module–Trait Correlations\n(cor, p-value, stars)"
)





labeledHeatmap(
  Matrix        = moduleTraitCor,
  xLabels       = colnames(traits),
  yLabels       = rownames(moduleTraitCor),
  ySymbols      = rownames(moduleTraitCor),
  colorLabels   = FALSE,
  colors        = colorRampPalette(c("blue", "white", "red"))(50),
  textMatrix    = textMatrix,
  setStdMargins = FALSE,    # use your par(mar=…) settings
  cex.text      = 0.6,
  zlim          = c(-1.1, 1.1),
  main          = "Module–Trait Correlations\n(cor, p-value, stars)"
)





#––––– Transparent PDF –––––#
cairo_pdf(
  filename = file.path("results", "03-WGCNA", "2.heatmap.pdf"),
  width    = 6,
  height   = 12,
  bg       = "transparent",
  family   = "Arial"
)
par(
  bg  = NA,              # plot background transparent
  mar = c(5, 12, 4, 2)    # bottom, left, top, right
)
labeledHeatmap(
  Matrix        = moduleTraitCor,
  xLabels       = colnames(traits),
  yLabels       = rownames(moduleTraitCor),
  ySymbols      = rownames(moduleTraitCor),
  colorLabels   = FALSE,
  colors        = colorRampPalette(c("blue", "white", "red"))(50),
  textMatrix    = textMatrix,
  setStdMargins = FALSE,    # respect our par(mar)
  cex.text      = 0.6,
  zlim          = c(-1, 1),
  main          = "Module–Trait Correlations\n(cor, p-value, stars)"
)
dev.off()

#––––– Transparent PNG –––––#
png(
  filename = file.path("results", "03-WGCNA", "2.heatmap.png"),
  width    = 6,
  height   = 12,
  units    = "in",
  res      = 300,          # adjust DPI as needed
  bg       = "transparent",
  
  family   = "Arial"
)
par(
  bg  = NA,
  mar = c(5, 12, 4, 2)
)
labeledHeatmap(
  Matrix        = moduleTraitCor,
  xLabels       = colnames(traits),
  yLabels       = rownames(moduleTraitCor),
  ySymbols      = rownames(moduleTraitCor),
  colorLabels   = FALSE,
  colors        = colorRampPalette(c("blue", "white", "red"))(50),
  textMatrix    = textMatrix,
  setStdMargins = FALSE,
  cex.text      = 0.6,
  zlim          = c(-1, 1),
  main          = "Module–Trait Correlations\n(cor, p-value, stars)"
)
dev.off()



```



```{r}
# # Required libraries
# library(clusterProfiler)
# library(org.Hs.eg.db)        # change if non-human
# library(dplyr)
# library(ggplot2)
# 
# # Your input: a data.frame with columns gene (symbol) and module
# # head(module_genes)
# #        gene module
# # 1     AADAT  black
# # 2   ABHD17A  black
# # ...
# # Make sure gene symbols match org.Hs.eg.db
# 
# # Base output directory
# base_dir <- "results/03-WGCNA/module_result"
# 
# # Get list of modules
# modules <- unique(module_genes$module)
# 
# for (mod in modules) {
#   # 1. Create module‐specific output dir
#   outdir <- file.path(base_dir, mod)
#   if (!dir.exists(outdir)) dir.create(outdir, recursive = TRUE)
#   
#   # 2. Extract gene list for this module
#   genes_sym <- module_genes %>%
#     filter(module == mod) %>%
#     pull(gene) %>%
#     unique()
#   
#   # 3. Map to Entrez IDs (required by clusterProfiler)
#   gene_df <- bitr(
#     genes_sym,
#     fromType   = "SYMBOL",
#     toType     = "ENTREZID",
#     OrgDb      = org.Hs.eg.db
#   )
#   entrez_ids <- gene_df$ENTREZID
#   
#   # 4. Run GO enrichment (Biological Process here)
#   ego <- enrichGO(
#     gene          = entrez_ids,
#     OrgDb         = org.Hs.eg.db,
#     ont           = "BP",
#     pAdjustMethod = "BH",
#     pvalueCutoff  = 0.05,
#     qvalueCutoff  = 0.2,
#     readable      = TRUE
#   )
#   
#   # 5. Save the full results table
#   res_tbl <- as.data.frame(ego)
#   write.csv(
#     res_tbl,
#     file = file.path(outdir, paste0(mod, "_GO_BP_enrichment.csv")),
#     row.names = FALSE
#   )
#   
#   # 6. Plot and save the top 20 terms as a dotplot
#   p <- dotplot(ego, showCategory = 20) +
#        ggtitle(paste0("Module ", mod, " — GO BP Enrichment"))
#   
#   ggsave(
#     filename = file.path(outdir, paste0(mod, "_GO_BP_dotplot.png")),
#     plot     = p,
#     width    = 8, height = 6
#   )
#   
#   message("Finished module: ", mod)
# }
# 
# 
# # make sure you have these loaded
# library(dplyr)
# 
# # your module⇢gene mapping
# # head(module_genes)
# #        gene module
# # 1     AADAT  black
# # 2   ABHD17A  black
# #  ...
# 
# # your gene_name_mapping object must already exist in your workspace
# 
# # Base directory for all module results
# base_dir <- "results/03-WGCNA/module_result"
# 
# # Get the unique modules
# modules <- unique(module_genes$module)
# 
# for (mod in modules) {
#   
#   # 1. Create the module-specific folder
#   outdir <- file.path(base_dir, mod)
#   if (!dir.exists(outdir)) {
#     dir.create(outdir, recursive = TRUE)
#   }
#   
#   # 2. Extract the gene symbols for this module
#   gene_list <- module_genes %>%
#     filter(module == mod) %>%
#     pull(gene) %>%
#     unique()
#   
#   # 3. Run your enrichment function
#   Enrichment_analysis(
#     gene_list         = gene_list,
#     result_folder     = outdir,
#     file_name         = paste0("module_", mod),
#     gene_name_mapping = gene_name_mapping
#     # flag defaults to "Up" – if you’d like a neutral color scale you can
#     # pass flag = "None" (or any string other than "Up"/"Down")
#   )
#   
#   message("Done enrichment for module: ", mod)
# }


```


```{r enrichment_analysis, fig.width=8, fig.height=4}
# ───────────────────────────────────────────────────────────────────────────────
# 1. Load libraries
# ───────────────────────────────────────────────────────────────────────────────
library(WGCNA)         # for corPvalueStudent()
library(gprofiler2)    # for gost() and map_genes()
library(dplyr)         # for data manipulation
library(ggplot2)       # for plotting

# ───────────────────────────────────────────────────────────────────────────────
# 2. Compute module–trait correlations & p‐values
# ───────────────────────────────────────────────────────────────────────────────
# (assumes MEs0 and condition_list are already in your environment)

# Convert your group factor into a numeric trait
traitNum <- ifelse(condition_list$group == "22q", 1, 0)
traits   <- data.frame(Phenotype = traitNum)

# Pearson correlations and Student’s p‐values
moduleTraitCor    <- cor(MEs0, traits, use = "p")
moduleTraitPvalue <- corPvalueStudent(moduleTraitCor,
                                      nSamples = nrow(MEs0))

# ───────────────────────────────────────────────────────────────────────────────
# 3. Identify “significant” modules (raw p < 0.05)
# ───────────────────────────────────────────────────────────────────────────────
pv       <- moduleTraitPvalue[, 1]
sig_mods <- rownames(moduleTraitPvalue)[pv < 0.05]

# ───────────────────────────────────────────────────────────────────────────────
# 4. Define the enrichment function (with 'parents'‐column fix)
# ───────────────────────────────────────────────────────────────────────────────
Enrichment_analysis <- function(gene_list,
                                result_folder,
                                file_name,
                                gene_name_mapping,
                                flag = "Up") {
  # 4.1 Run g:Profiler
  enrichment <- gost(
    query             = gene_list,
    organism          = "hsapiens",
    correction_method = "fdr",
    evcodes           = TRUE
  )
  
  # 4.2 Drop the problematic 'parents' column
  result <- enrichment$result %>%
    dplyr::select(-parents)
  
  # 4.3 Map back to your original gene names (using gprofiler2 metadata)
  result <- map_genes(
    result,
    enrichment$meta$genes_metadata$query$query_1$mapping,
    gene_name_mapping
  )
  
  # 4.4 Create a combined term label
  result <- result %>%
    mutate(term = sprintf("%s - %s", source, term_name))
  
  # 4.5 Save the full result table
  write.csv(
    result,
    file.path(result_folder, paste0(file_name, ".csv")),
    row.names = FALSE
  )
  message("Saved full enrichment for: ", file_name)
  
  # ─────────────────────────────────────────────────────────────────────────────
  # 4.6 Plot top 30 pathways (all sources)
  # ─────────────────────────────────────────────────────────────────────────────
  top30 <- result %>%
    dplyr::select(term, p_value) %>%
    arrange(p_value) %>%
    slice_head(n = 30) %>%
    mutate(term = factor(term, levels = rev(term)))
  
  p <- ggplot(top30, aes(x = term, y = -log10(p_value), fill = -log10(p_value))) +
    geom_bar(stat = "identity") +
    { if (flag == "Up") {
      scale_fill_gradient(high = "#a50f15", low = "#fc9272")
    } else if (flag == "Down") {
      scale_fill_gradient(low = "#56B1F7", high = "#132B43")
    } else {
      scale_fill_manual(values = rep("lightblue", nrow(top30)))
    }
    } +
    ylab("-log10(FDR)") +
    xlab("") +
    ggtitle("Top 30 Pathways (Sorted by FDR)") +
    coord_flip() +
    theme_classic(base_size = 14) + theme(
    axis.text   = element_text(color = "black"),
    axis.title  = element_text(color = "black"),
    legend.text = element_text(color = "black"),
    legend.title= element_text(color = "black")
  )+
    theme(
      text       = element_text(family = "Arial", size = 14, colour = "black"),
      plot.title = element_text(face = "bold", hjust = 0.5),
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background  = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent", colour = NA)
    ) +
    geom_hline(yintercept = -log10(0.05), linetype = "dashed")
  
  ggsave(
    filename = file.path(result_folder, paste0(file_name, "_top30_pathways.png")),
    plot     = p, width = 12, height = 8, dpi = 300,
    bg       = "transparent"
  )
  ggsave(
    filename = file.path(result_folder, paste0(file_name, "_top30_pathways.pdf")),
    plot     = p, width = 12, height = 8,
    bg       = "transparent"
  )

  message("Plotted top 30 pathways for: ", file_name)
  
  # ─────────────────────────────────────────────────────────────────────────────
  # 4.7 Subset to GO:BP, save & plot top 30
  # ─────────────────────────────────────────────────────────────────────────────
  go_bp <- result %>% filter(source == "GO:BP")
  write.csv(
    go_bp,
    file.path(result_folder, paste0(file_name, "_GO_BP.csv")),
    row.names = FALSE
  )
  
  top30_bp <- go_bp %>%
    dplyr::select(term, p_value) %>%
    arrange(p_value) %>%
    slice_head(n = 30) %>%
    mutate(term = factor(term, levels = rev(term)))
  
  p2 <- ggplot(top30_bp, aes(x = term, y = -log10(p_value), fill = -log10(p_value))) +
    geom_bar(stat = "identity") +
    { if (flag == "Up") {
      scale_fill_gradient(high = "#a50f15", low = "#fc9272")
    } else if (flag == "Down") {
      scale_fill_gradient(low = "#56B1F7", high = "#132B43")
    } else {
      scale_fill_manual(values = rep("lightblue", nrow(top30_bp)))
    }
    } +
    ylab("-log10(FDR)") +
    xlab("") +
    ggtitle("Top 30 GO:BP Pathways (Sorted by FDR)") +
    coord_flip() +
    theme_classic(base_size = 14) +
    theme(
      text       = element_text(family = "Arial", size = 14, colour = "black"),
      plot.title = element_text(face = "bold", hjust = 0.5)
    ) + theme(
    axis.text   = element_text(color = "black"),
    axis.title  = element_text(color = "black"),
    legend.text = element_text(color = "black"),
    legend.title= element_text(color = "black"),
    panel.background = element_rect(fill = "transparent", colour = NA),
    plot.background  = element_rect(fill = "transparent", colour = NA),
    legend.background = element_rect(fill = "transparent", colour = NA)
  )+
    geom_hline(yintercept = -log10(0.05), linetype = "dashed")
  
  ggsave(file.path(result_folder,
                   paste0(file_name, "_top30_pathways_GOBP.png")),
         p2, width = 12, height = 8, dpi = 300,
  bg       = "transparent")
  ggsave(file.path(result_folder,
                   paste0(file_name, "_top30_pathways_GOBP.pdf")),
         p2, width = 12, height = 8,
  bg       = "transparent")
  message("Plotted GO:BP top 30 for: ", file_name)
}

# ───────────────────────────────────────────────────────────────────────────────
# 5. Loop over significant modules
# ───────────────────────────────────────────────────────────────────────────────
base_dir <- "results/03-WGCNA/module_result"

for (mod in sig_mods) {
  print(sprintf("Processing module: %s", mod))
  outdir <- file.path(base_dir, mod)
  if (!dir.exists(outdir)) dir.create(outdir, recursive = TRUE)
  
  # 2) get the correlation value for this module → phenotype
  cor_val <- moduleTraitCor[mod, "Phenotype"]      # or [,1] if 'Phenotype' is column 1
  flag    <- if (cor_val > 0) "Up" else "Down"
  
  mod <- gsub("^ME", "", mod)  # remove "ME" prefix from module name
  
  gene_list <- module_genes %>%
    filter(module == mod) %>%
    pull(gene) %>%
    unique()
  
  unique(module_genes$module)
  

  
  Enrichment_analysis(
    gene_list         = gene_list,
    result_folder     = outdir,
    file_name         = paste0("module_", mod),
    gene_name_mapping = gene_name_mapping,
    flag              =   flag 
  )
  
  message("Enrichment done for module: ", mod)
}

```



\newpage
# Session information
```{r}
sessionInfo()
```

